---
title: "Data manipulation"
---

```{r eval = TRUE,  message=F, include=F, warning=F, purl=F, results="hide"}
knitr::purl('manipulation.Rmd', documentation = F)
```

```{r echo=FALSE, purl=F}
xfun::embed_file('manipulation.Rmd')
```

```{r echo=FALSE, purl=F}
xfun::embed_file('manipulation.R')
```


Manipulating a data set involves modifying the data to make them easier to read and/or to make the entire data set more organized for the following analyses. Commonly, we manipulate data prior their visualization and/or analysis. It is often used and combined with the term ‘data exploration’ which involves organizing data with an available sets of variables. At times, the data collection process done by machines involves a lot of errors and inaccuracies in reading. Data manipulation is used to remove these inaccuracies and make data more accurate and precise.

# Dataset

The package `datasets` provides data sets. It is part of the basics installation of R and you don't have to install the package (if you accept default option at R installation). Those data sets are used in many examples through help in R. Among those data sets, `iris` (`?iris`) provides measurements of the variables sepal length and width and petal length and width, respectively, for 50 flowers from each of 3 species of iris. The `Species` is a categorical variable with three levels: *Iris setosa*, *versicolor*, and *virginica*. Flower measurements are quantitative and continous  variables, expressed in centimeters.   

```{r,  eval=T}
rm(list=ls()) # clean memory
library (datasets) # load package
data(iris) # import dataset
head (iris) # visualize 'head' dataset
# 'data' automatically  creates the object 'iris'
```

Summarize this dataset: 

```{r,  eval=T}
summary(iris) #  object summary
```

You can examine the `class` (type) of the object you are manipulating. Examining its structure `str` is essential to understand the variables you are dealing with.

```{r,  eval=T}
str(iris) # examine the structure of the object
```

The object `iris` is a `data.frame` and is composed of 5 variables (`Sepal.Length`, `Sepal.Width`, `Petal.Length`, `Petal.Width`) and 150 observations. Four variables are numeric, one is a factor with 3 levels (`Species`). Other object types and structures also exist (integer, matrix, etc.) but more on this later (see [Types & Structures](./types.html)).

Summarizing your data set and examining its structure allow you to spot a possible error in your data set (including format errors that may have been caused at the importation of data set, remember `reef_fish.txt`). You can fix format error using `read.table` or others. Alternatively, you can `fix` it to open a spreadsheet:

```{r,  eval=FALSE}
fix(iris) # spreadsheet
```

# Selection

## Column and row

Let's assume an online data set documenting the height and shoe size of 10 students in Taiwan (we could make our own data set, but we want to import one from previous year and available online):

<div style="overflow: auto !important;>

```{r,  eval=F}
students<-read.table('https://www.dipintothereef.com/uploads/3/7/3/5/37359245/students.txt',header=T, sep="\t", dec='.') # read data set from url
str(students) 
```

```{r,  eval=T, echo=F}
students<-read.table('https://www.dipintothereef.com/uploads/3/7/3/5/37359245/students.txt',header=T, sep="\t", dec='.') # read data set from url
str(students) 
```


`r knitr::asis_output("\U1F449")` *Comment on the structure of this data set*

A single column (height) is needed for further processing. This is accomplished by using the `$` operator. A column can be selected by putting together on the same line the:

- the name of the object `students`
- the operator `$`
- the name of the column, either `height`, `shoesize`, `gender`, `population`    

Here to select the column `height`, we can simply use `students$height`. Note that the same column can be extracted using  **subscript**  which is denoted by the **square brackets** after the object name such as `object [no. row, no. column]`. Therefore to select the first column of our object `students`,  one can also used: `students[,1]`

```{r,  eval=T}
# students[,1]
students$height
```

Extracting a row can only be accomplished with a subscript. Therefore, to extract the first row of the dataframe:

```{r,  eval=T}
students[1,]
```

Now  that you understand how to select column and row, you will understand the following chunk of code: 

```{r,  eval=FALSE}
students[1,1] # element in the first row, first column
students$height[1] # first element in our vector height
```

<span style="color: red"> **Remember:** `object[no. row, no. column]`</span>

## Subset

The purpose of subsetting is to  extract information following a given condition. Such, we only want to select  `female` students in the `students` data set. Such as selection can be made by *filtering* `students` using `gender` information. Let's examine the logic behind:

1. Add a condition and check whether or not the observation (row) is for female students:

```{r,  eval=T}
students$gender=="female" # condition
```

2. Let's put this condition into a *filter*. Let's call it `f` as it stands for `female`: 

```{r,  eval=T}
f<-students$gender=="female" # filter
``` 

3. Apply this *filter* `f` using subscript. We want to extract/select only the **rows** that contain data for `female` students.Resulting data set is put into a new object call `females`:

```{r,  eval=T}
females<-students[f,] # selection
females
```

Our subset is constituted of a selection of the five female students. The original `students` data set had ten observations, and I can check this change calling for the dimensions of the new object `females` and using `dim(females)`. The new data set is a `data.frame` of 5 observations x 4 columns. Column names `colnames(females)` remained unchanged, but I have now only 5 names for the rows `rownames(females)`. The rows can be renamed by  combining `c` the names of the female students into a **vector**.


```{r,  eval=T}
rownames(females)<-c('Vanessa', 'Vicky', 'Michelle', 'Joyce', 'Victoria')
females
``` 

<p class="comment">
**Practice 2.1** Using the `iris` data set, create three smaller data sets corresponding to the three species designated in the column `Species`. Do not look at the solution before trying.
</p>

```{r, code_folding = 'Show Solution',  eval=FALSE}
# NOT RUN:  levels(iris$Species)
setf<-iris$Species=="setosa"
setosa<-iris[setf,]
verf<-iris$Species=="versicolor"
versicolor<-iris[verf,] 
virf<-iris$Species=="virginica"
virginica<-iris[virf,] 
```

## Sample

The `sample` function can be used to select one of you to illustrate the above subsets (see [Home](./index.html)). More generally, the function `sample` is used to generate a sample of a given size and extracted from a vector or a data set; either with or without replacement.

The basic syntax of the `sample` function is as follows:

```{r class.source = "fold-show",  eval=FALSE}
sample(data, size, replace = FALSE, prob = NULL)
# check in detail source code 
# View(sample) OR getAnywhere(sample()) 
```

*You can check in details what the function is exactly doing by checking its source code: `View(sample)` OR `getAnywhere(sample())`. It is important if you wanna  carry a modification in an existing function.*  
Therefore, for a random selection of two individuals in our `females` data set, the steps are: (1) define the size of the targeted vector; (2) create filter with individual(s) randomly selected; (3) apply the filter on the data set: 


```{r,  eval=T}
# NOT RUN nrow(females)
1:nrow(females) # create a vector from 1 to the number of row in 'females'
sf<-sample(1:nrow(females), 2) # filter with two randomly selected female students
sf # the selection
females[sf,] # apply the filter on our original data set
```

# Sorting

Sorting a data frame is done by combining function `order` with a subscript. Exactly as we did earlier while subsetting, we first need to create a vector that will tell into which order the rows of the table need to be arranged. Let's `order` the `students` by increasing `height`:

```{r,  eval=FALSE}
# NOT RUN, check data set ind no.1 should move at position no. 5
ind1<-order (students$height) # create a vector with order
students [ind1,] # sorting data set with  with filter
# ind no.1 moved at position no. 5
```

Both can be combined into on line to be more concise (and elegant). Yes we can speak about the beauty of R script:

```{r,  eval=T}
students[order(students$height),]
```

You can now guess what the following vectors will do: 

```{r,  eval=FALSE}
ind2<-order(-students$height)
```

It is the equivalent of `ind3<-order(students$height,decreasing =T)`


# Recoding

Recoding is replacing values with others. Values in a variable can be recoded using the command `ifelse`. Let's recode the `gender` variable in `students` with color names. We will replace  `male` information with `blue` and `female` information with `red`. Note that here, we will use the operator `==` which means strictly 'equal to'. Basically, to set a given condition and impose a task if this condition is true and alternative task if this condition is not true:

- Conditional element selection

```{r,  eval=T}
# If `condition` then give 'blue'. If not then gives 'red'.
colors<-ifelse(students$gender=='male', 'blue','red') 
colors
```

- Create a new column

```{r,  eval=T}
students$colors<-ifelse(students$gender=='male','blue','red') # create a new column
```

- Replace an existing column

```{r,  eval=F}
students$gender<-ifelse(students$gender=='male', 'blue','red') # replace an existing column
```

More operators, which can be combined together:

- `==` equal to 

- `>=` equal to or greater than

- `<=` equal to or less than

- `!=` not equal to

- `&` and

- `|` or

Examples:

```{r,  eval=F}
students$height <= 165 # students shorter or equal to 165 cm
students$shoesize < 37 # students with shoes size smaller than 37
students$height <= 165 & students$shoesize < 37 # students equal or shoter than 165 and shoes size smaller than 37
# NOT RUN: students$dual.cond<-ifelse(students$height<=165 & students$shoesize<37,'blue','red') 
```

Let's finally put all this in practice. 

<p class="comment">
**Practice 2.2** Using the `iris` dataset create a new variable with flower colors. Iris setosa is purple.Iris versicolor is blue. Iris virginica is pink. Sort individuals by decreasing `Sepal.Width`. What can you hypothesize on the size of sepal for these three species. Get back the data set for the species having the smallest sepal width. Delete the variable `color` in this subset</span>* 
</p>


```{r, code_folding = 'Show Solution',  eval=FALSE}
# ask for password
```

<form id="form" onsubmit="return false;">
  <input type="password" id="userInput" />
  <input type="submit" onclick="othername();" />
</form>

<p id="hidden_clue"></p>


<script>
function othername() {
    var pass1 = 12345;
    var input = document.getElementById("userInput").value;
    if (input == pass1) {
        document.getElementById("hidden_clue").textContent = "iris$color<-ifelse(iris$Species=='setosa','purple', ifelse(iris$Species=='versicolor','blue','pink')) THEN iris[order(iris$Sepal.Width, decreasing = T),]";
    }
}
</script>

`r knitr::asis_output("\U1F449")` *To delete a variable, this variable will received `NULL`*

Those are functions in R Base Package. See also `?subset`, `?with`, `?within`, etc. Other packages such as `data.table`, `reshape2`, `dplyr`, `tidyr`, etc. offer useful functionality for manipulating data set. Let's see later for the `tidyverse` universe. 