---
title: "Practice 7.4"
author: "Shin MJ"
date: "2022-11-06"
output: html_document
---
##**T-test source code**

```{r, echo= TRUE}
data("rairuoho")

tttest<- function (x, y = NULL, alternative = c("two.sided", "less", "greater"), 
          mu = 0, paired = FALSE, var.equal = FALSE, conf.level = 0.95, 
          ...) 
{
  alternative <- match.arg(alternative)
  if (!missing(mu) && (length(mu) != 1 || is.na(mu))) 
    stop("'mu' must be a single number")
  if (!missing(conf.level) && (length(conf.level) != 1 || !is.finite(conf.level) || 
                               conf.level < 0 || conf.level > 1)) 
    stop("'conf.level' must be a single number between 0 and 1")
  if (!is.null(y)) {
    dname <- paste(deparse1(substitute(x)), "and", deparse1(substitute(y)))
    if (paired) 
      xok <- yok <- complete.cases(x, y)
    else {
      yok <- !is.na(y)
      xok <- !is.na(x)
    }
    y <- y[yok]
  }
  else {
    dname <- deparse1(substitute(x))
    if (paired) 
      stop("'y' is missing for paired test")
    xok <- !is.na(x)
    yok <- NULL
  }
  x <- x[xok]
  if (paired) {
    x <- x - y
    y <- NULL
  }
  nx <- length(x)
  mx <- mean(x)
  vx <- var(x)
  if (is.null(y)) {
    if (nx < 2) 
      stop("not enough 'x' observations")
    df <- nx - 1
    stderr <- sqrt(vx/nx)
    tstat <- (mx - mu)/stderr
    method <- if (paired) 
      "Paired t-test"
    else "One Sample t-test"
    estimate <- setNames(mx, if (paired) 
      "mean difference"
      else "mean of x")
  }
  else {
    ny <- length(y)
    my <- mean(y)
    vy <- var(y)
    method <- paste(if (!var.equal) 
      "Welch", "Two Sample t-test")
    estimate <- c(mx, my)
    names(estimate) <- c("mean of x", "mean of y")
    if (var.equal) {
      df <- nx + ny - 2
      v <- 0
      if (nx > 1) 
        v <- v + (nx - 1) * vx
      if (ny > 1) 
        v <- v + (ny - 1) * vy
      v <- v/df
      stderr <- sqrt(v * (1/nx + 1/ny))
    }
    else {
      stderrx <- sqrt(vx/nx)
      stderry <- sqrt(vy/ny)
      stderr <- sqrt(stderrx^2 + stderry^2)
      df <- stderr^4/(stderrx^4/(nx - 1) + stderry^4/(ny - 
                                                        1))
    }
    if (stderr < 10 * .Machine$double.eps * max(abs(mx), 
                                                abs(my))) 
      stop("data are essentially constant")
    tstat <- (mx - my - mu)/stderr
  }
  if (alternative == "less") {
    pval <- pt(tstat, df)
    cint <- c(-Inf, tstat + qt(conf.level, df))
  }
  else if (alternative == "greater") {
    pval <- pt(tstat, df, lower.tail = FALSE)
    cint <- c(tstat - qt(conf.level, df), Inf)
  }
  else {
    pval <- 2 * pt(-abs(tstat), df)
    alpha <- 1 - conf.level
    cint <- qt(1 - alpha/2, df)
    cint <- tstat + c(-cint, cint)
  }
  cint <- mu + cint * stderr
  names(tstat) <- "t"
  names(df) <- "df"
  names(mu) <- if (paired) 
    "mean difference"
  else if (!is.null(y)) 
    "difference in means"
  else "mean"
  attr(cint, "conf.level") <- conf.level
  rval <- list(statistic = tstat, parameter = df, p.value = pval, 
               conf.int = cint, estimate = estimate, null.value = mu, 
               stderr = stderr, alternative = alternative, method = method, 
               data.name = dname)
  class(rval) <- "htest"
  rval
}

```

``` {r}
n3 <- rairuoho[rairuoho$treatment =="nutrient", ]$day3
n4 <- rairuoho[rairuoho$treatment =="nutrient", ]$day4
n5 <- rairuoho[rairuoho$treatment =="nutrient", ]$day5
n6 <- rairuoho[rairuoho$treatment =="nutrient", ]$day6
n7 <- rairuoho[rairuoho$treatment =="nutrient", ]$day7
n8 <- rairuoho[rairuoho$treatment =="nutrient", ]$day8
w3 <- rairuoho[rairuoho$treatment =="water",]$day3
w4 <- rairuoho[rairuoho$treatment =="water",]$day4
w5 <- rairuoho[rairuoho$treatment =="water",]$day5
w6 <- rairuoho[rairuoho$treatment =="water",]$day6
w7 <- rairuoho[rairuoho$treatment =="water",]$day7
w8 <- rairuoho[rairuoho$treatment =="water",]$day8

n3w3 <-tttest(n3, w3, paired = FALSE, alternative ="two.sided", var.equal= FALSE)
n3w3
```

##Simple code for T-test 
Since I am getting different results, I know there is something wrong with the code. But my problem is that I cannot identify what is actually wrong. **SEND HELP**

*the formula maybe?
*the code itself?
**I simply find it quite difficult to connect both statistics and coding. ;__;


``` {r, echo= TRUE}
myttest<- function (x,y, alternative = c("two.sided", "less", "greater"),
                    paired = FALSE, var.equal = FALSE, conf.level= 0.95) {
  nx<-length(x)
  ny<-length(y)
  mx<-mean(x)
  my<-mean(y)
  df<-nx + ny - 2 
  s2<-((sum(x-mx)^2)+(sum(y-my)^2)/(nx+ny-2))
  tstat<-(mx-my)/(sqrt((s2/nx)+(s2/ny)))
  wtstat<- (mx+my)/(sqrt(((sd(x))^2/nx)/(sd(y)/ny)^2))
  df1<- (((sd(x))^2/nx)/(sd(y)/ny)^2/ny)
    ((sd(x))^4/((nx)^2-(nx - 1))+
    (sd(y)^4)/((ny)^2-(ny - 1)))
{
  alternative <- match.arg(alternative)
  if (!missing(conf.level) && (length(conf.level) != 1 || !is.finite(conf.level) || 
                               conf.level < 0 || conf.level > 1)) 
    stop("'conf.level' must be a single number between 0 and 1")
  if (!is.null(y)) {
    dname <- paste(deparse1(substitute(x)), "and", deparse1(substitute(y)))
    if (paired) 
      xok <- yok <- complete.cases(x, y)
    else {
      yok <- !is.na(y)
      xok <- !is.na(x)
    }
    y <- y[yok]
  }
  else {
    dname <- deparse1(substitute(x))
    if (paired) 
      stop("'y' is missing for paired test")
    xok <- !is.na(x)
    yok <- NULL
  }
  x <- x[xok]
  if (paired) {
    x <- x - y
    y <- NULL
  }
}
{
  if (alternative == "less") {
    pval <- pt(tstat, df)
    cint <- c(-Inf, tstat + qt(conf.level, df))
  }
  else if (alternative == "greater") {
    pval <- pt(tstat, df, lower.tail = FALSE)
    cint <- c(tstat - qt(conf.level, df), Inf)
  }
  else {
    pval <- 2 * pt(-abs(tstat), df)
    alpha <- 1 - conf.level
    cint <- qt(1 - alpha/2, df)
    cint <- tstat + c(-cint, cint)
  }
  names(tstat) <- "t"
  names(df) <- "df"
  attr(cint, "conf.level") <- conf.level
  rval <- list(statistic = tstat, parameter = df, p.value = pval, 
               conf.int = cint, stderr = stderr, alternative = alternative, 
               data.name = dname)
  class(rval) <- "htest"
  rval
}
}


n3w3<- myttest(n3,w3, paired = FALSE, alternative ="two.sided", var.equal= FALSE)
n3w3


